---
title: "Inferring Arithmetic Skill from Speed and Accuracy."
abstract: |
  People routinely infer others' competence under uncertainty, often relying on cues such as task difficulty and past accuracy. An emerging body of research suggests that people approximate Bayesian inference when doing so. We extend these results by testing whether people can infer others' numerical ability in a way that is consistent wih th a rational Bayesian model. In Study 1, we find that participants accurately predict the arithmetic performance of another individual from information about their past performance. Computational modeling shows that participants' inferences are better described by Bayesian processes than by plausible heuristics. Study 2 introduces a modified paradigm, in which participants are told about both past performance and time taken to solve problems. We find that, although participants are quite accurate in their predictions, they do not seem to take into account information about speed.

output:
  papaja::apa6_pdf:
    keep_tex: true
  papaja::apa6_docx:
always_allow_html: true
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra") # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300) # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center") # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.height = 5.25,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "figures/"
  # cache.path = "output/_cache/",
  # fig.process = function(x) {  # Remove "-1" from figure names
  #   x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  #   if (file.rename(x, x2)) x2 else x
  # },
  # options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90, # Output width
  dplyr.summarise.inform = FALSE, # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "", # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE, # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format, # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain" # Don't use siunitx
)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(mlogit) # For multinomial logistic regression analysis
library(flextable) # For creating formatted tables (Word output)
library(kableExtra) # For creating formatted tables (PDF output)
library(officer) # For knitting formatted tables to word
library(papaja) # For APA-style formatting
library(lmerTest) # For linear mixed-effects models with p-values
library(broom) # For tidying model outputs
library(broom.mixed) # For tidying mixed effects model outputs
library(wesanderson) # For color palettes inspired by Wes Anderson films
library(patchwork) # For combining multiple ggplots
library(here) # For project-relative file paths
library(ggh4x) # For extended ggplot2 functionality
library(car) # For Anova function
library(ggrepel) # For non-overlapping point labels
```

```{r functions}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 2)

# get model outputs ready for inline text reporting
clean_lm <- function(model_output) {
  result <- tidy(model_output, conf.int = TRUE) %>%
    # filter(effect == "fixed") %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      # if there is an interaction (recognized by ":"), name it just interaction
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, statistic, p.value) %>%
    split(.$clean_term)

  model_fit <- glance(model_output) %>%
    select(adj.r.squared, df.residual, df, statistic, p.value) %>%
    rounded_numbers() %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit

  return(result)
}

clean_lmer <- function(model_output) {
  result <- broom.mixed::tidy(model_output, conf.int = TRUE) %>%
    filter(effect == "fixed") %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      # if there is an interaction (recognized by ":"), name it just interaction
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, any_of("df"), statistic, p.value) %>%
    split(.$clean_term)

  return(result)
}

# clean data frame results for paired t-test
clean_t_test <- function(t.test) {
  t.test %>%
    tidy() %>%
    select(-c(method, alternative)) %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })
}

clean_mlogit <- function(model_output) {
  # Get the summary of the mlogit model, which includes a coefficient table
  summ <- summary(model_output)

  # Convert the coefficient matrix to a data frame with row names as a column
  coef_table <- as.data.frame(summ$CoefTable)
  coef_table <- tibble::rownames_to_column(coef_table, var = "term")

  # Rename columns to common names (make sure these match your summary output)
  names(coef_table) <- c("term", "estimate", "std.error", "z.value", "p.value")

  # Process the coefficient table
  coef_table <- coef_table %>%
    mutate(
      # Format p-values per APA guidelines
      p.value = case_when(
        p.value < 0.001 ~ "< .001",
        TRUE ~ paste0("= ", sprintf("%.3f", p.value))
      ),
      # Calculate confidence intervals (normal approximation)
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      ci = glue::glue("[{round(conf.low, 2)}, {round(conf.high, 2)}]"),
      # Rename intercept and collapse any interaction terms
      term = ifelse(term == "(Intercept)", "intercept", term),
      term = ifelse(str_detect(term, ":"), "interaction", term)
    ) %>%
    select(term, estimate, std.error, ci, z.value, p.value) %>%
    rounded_numbers()

  # Split the table by term so each coefficient gets its own list element
  result <- split(coef_table, coef_table$term)

  # Create a model fit summary.
  # Here we include logLik, AIC, BIC, and the number of iterations (if available)
  model_fit <- data.frame(
    logLik = as.numeric(model_output$logLik),
    AIC = AIC(model_output),
    BIC = -2 * as.numeric(model_output$logLik) + attr(logLik(model_output), "df") * log(length(model_output$fitted.values)),
    nb.iter = model_output$est.stat$nb.iter
  )
  model_fit <- rounded_numbers(model_fit) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit

  return(result)
}

custom_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    ## ---- facet-label styling ----
    strip.text = element_text(colour = "black"), # text colour
    strip.background = element_rect(
      fill = "white", # inside of the strip
      colour = "black", # border line
      linewidth = .8
    ),

    ## ggplot2 ≥ 3.4 splits strips into x- and y-orientations;
    ## setting these keeps the look identical for both directions
    strip.background.x = element_rect(fill = "white", colour = "black", linewidth = .8),
    strip.background.y = element_rect(fill = "white", colour = "black", linewidth = .8)
  )
# panel.background = element_rect(fill = "white", colour = NA),
# panel.border     = element_rect(fill = NA, colour = "black", linewidth = .4))

# Function to clean p-values for inline reporting
clean_p <- function(p_value) {
  if (p_value < 0.001) {
    return("< .001")
  } else {
    return(sprintf("%.3f", p_value))
  }
}


# Theme for plots in slides:
custom_theme_dark <- theme_minimal() +
  theme(
    text = element_text(family = "serif", color = "white", size = 14),
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid.major = element_line(color = "grey30"),
    panel.grid.minor = element_line(color = "grey20"),
    axis.text = element_text(color = "white", size = 12),
    axis.title = element_text(color = "white", size = 14),
    plot.title = element_text(color = "white", size = 18, face = "bold"),
    legend.background = element_rect(fill = "black", color = NA),
    legend.key = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white", size = 12),
    legend.title = element_text(color = "white", size = 14),
    strip.text = element_text(color = "white", size = 14)
  )
```

```{r S1-load-data}
S1_long_data   <- read_csv(here::here("study_1", "data", "clean", "S1_long_data.csv"))
S1_prob_df     <- read_csv(here::here("study_1", "data", "clean", "S1_prob_df.csv"))
S1_difficulty  <- read_csv(here::here("study_1", "data", "clean", "S1_difficulty.csv"))

source(here::here("pre-registered code", "pre-reg_model_implementation.R"))

S1_fit_results <- read_csv(here::here("study_1", "results", "S1_fit_results.csv"))
```

```{r S1-analyses}
S1_long_data <- left_join(
  S1_long_data, S1_prob_df,
  by = c("question_observed", "success_observed", "question_evaluated")
)

# Bottom 30% subsample (robustness)
n_participants <- S1_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S1_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S1_long_data_worst <- S1_long_data %>% filter(prolific_pid %in% worst_ids)

# Summarized data (one row per observed-new question pair x success/failure)
S1_summarized_data <- S1_long_data %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    judged_prob = mean(human_evaluation, na.rm = TRUE),
    observed_q_perceived_difficulty_unbounded = mean(observed_q_perceived_difficulty_unbounded),
    new_q_perceived_difficulty_unbounded = mean(new_q_perceived_difficulty_unbounded),
    true_prob = mean(true_conditional_prob),
    .groups = "drop"
  ) %>%
  filter(question_observed != question_evaluated)

# ---- Bayesian / heuristic model predictions ----
S1_MAIN_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$LLMax + 4 * log(nrow(S1_long_data))
S1_NULL_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL", ]$LLMax + 1 * log(nrow(S1_long_data))
S1_NULL_2_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$LLMax + 2 * log(nrow(S1_long_data))

delta <- 0.005
COMP  <- seq(-4, 4, by = delta)

mu    <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$mu
sigma <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$sigma
noise <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$noise
beta  <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$beta

S1_prior_COMP <- dnorm(COMP, mu, sigma)
S1_prior_COMP <- S1_prior_COMP / sum(S1_prior_COMP * delta)

S1_summarized_data$prob_bayes <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_prior_COMP, noise = noise, beta = beta)
)

beta <- S1_fit_results[S1_fit_results$model == "S1_NULL", ]$beta
S1_NULL_COMP <- rep(1, length(COMP))
S1_NULL_COMP <- S1_NULL_COMP / sum(S1_NULL_COMP * delta)

S1_summarized_data$prob_null <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_NULL_COMP, noise = 0, beta = beta)
)

beta       <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$beta
difference <- beta <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$diff

S1_summarized_data$prob_null_2 <- mapply(
  FUN = function(obs, new, success_observed) {
    if (success_observed == 1) {
      soft_max(ifelse(obs + difference >= new, 1, 0), beta = beta)
    } else {
      soft_max(ifelse(obs - difference >= new, 1, 0), beta = beta)
    }
  },
  obs = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success_observed = S1_summarized_data$success_observed
)

# ---- Confirmatory models ----
data_h1 <- S1_long_data %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty_indiv) %>%
  distinct()
model_h1 <- lmer(scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1)

model_h2 <- lm(scale(judged_prob) ~ scale(true_prob), data = S1_summarized_data)

model_bayes <- lm(scale(judged_prob) ~ scale(prob_bayes), data = S1_summarized_data)
model_null1 <- lm(scale(judged_prob) ~ scale(prob_null),  data = S1_summarized_data)
model_null2 <- lm(scale(judged_prob) ~ scale(prob_null_2), data = S1_summarized_data)

# ---- Robustness (bottom 30%) ----
data_h1_worst <- S1_long_data_worst %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty) %>%
  distinct()
model_h1m <- lmer(scale(observed_q_perceived_difficulty) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1_worst)

data_h2m <- S1_long_data_worst %>%
  group_by(question_observed, question_evaluated, success_observed) %>%
  summarise(
    true_prob   = mean(true_conditional_prob),
    judged_prob = mean(human_evaluation),
    .groups = "drop"
  )
model_h2m <- lm(scale(judged_prob) ~ scale(true_prob), data_h2m)

S1_results <- list(
  clean_model_h1    = clean_lmer(model_h1),
  clean_model_h2    = clean_lm(model_h2),
  clean_model_bayes = clean_lm(model_bayes),
  clean_model_null1 = clean_lm(model_null1),
  clean_model_null2 = clean_lm(model_null2),
  clean_model_h1m   = clean_lmer(model_h1m),
  clean_model_h2m   = clean_lm(model_h2m)
)
```

# Study 1

**H1 (Difficulty judgments):** $\beta$ = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$p.value`.

**H2 (Competence inferences from a single outcome):** $\beta$ = `r S1_results$clean_model_h2$true_prob$estimate`, t(`r S1_results$clean_model_h2$model_fit$df.residual`) = `r S1_results$clean_model_h2$true_prob$statistic`, p `r S1_results$clean_model_h2$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2$model_fit$adj.r.squared`.

**H3--H4 (Model accuracy and comparison):**

- Main (Bayesian) Model: $\beta$ = `r S1_results$clean_model_bayes$prob_bayes$estimate`, t(`r S1_results$clean_model_bayes$model_fit$df.residual`) = `r S1_results$clean_model_bayes$prob_bayes$statistic`, p `r S1_results$clean_model_bayes$prob_bayes$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_bayes$model_fit$adj.r.squared`.
- Threshold: Adjusted $R^2$ = `r S1_results$clean_model_null1$model_fit$adj.r.squared`.
- Anchor: Adjusted $R^2$ = `r S1_results$clean_model_null2$model_fit$adj.r.squared`.
- BIC: $BIC_{\text{Bayes}}$ = `r round(S1_MAIN_BIC, 2)`; $BIC_{\text{Threshold}}$ = `r round(S1_NULL_BIC, 2)`; $BIC_{\text{Anchor}}$ = `r round(S1_NULL_2_BIC, 2)`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S1_results$clean_model_h2m$true_prob$estimate`, t(`r S1_results$clean_model_h2m$model_fit$df.residual`) = `r S1_results$clean_model_h2m$true_prob$statistic`, p `r S1_results$clean_model_h2m$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2m$model_fit$adj.r.squared`.

```{r S1-fig-behav, fig.cap = "(\\textbf{A}) Each question's estimated difficulty as a function of its true difficulty. Numbers refer to the question ID (see Materials). The red dashed line represents a perfectly accurate estimation. (\\textbf{B}) Participants' estimated probability of answering a new question correctly, conditioned on the performance on the observed question, as a function of true conditional probabilities. (\\textbf{C}) The Bayesian model's estimated conditional probabilities as a function of true conditional probabilities. For (\\textbf{B}) and (\\textbf{C}), each data point corresponds to the average prediction for an observed-new question pair. The red line represents a perfectly accurate prediction, while the dark line represents the fit of the linear model.", fig.width = 7, fig.height = 4}
A <- S1_difficulty %>%
  ggplot(aes(x = objective_difficulty, y = perceived_difficulty)) +
  geom_point(color = "#E4A70A") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  geom_label_repel(
    aes(label = question),
    color = "#E4A70A",
    size = 2.5, show.legend = FALSE, seed = 42,
    max.overlaps = Inf, box.padding = 0.4, point.padding = 0.15,
    label.padding = grid::unit(0.12, "lines"),
    label.size = 0.3, label.r = grid::unit(0.15, "lines"),
    fill = scales::alpha("white", 0.7),
    force = 3, force_pull = 0.2, max.iter = 50000,
    min.segment.length = 0, segment.color = "grey60", segment.size = 0.3
  ) +
  xlim(0, 1) + ylim(0, 1) +
  labs(x = "True difficulty", y = "Estimated difficulty", title = "Participants'\nestimations") +
  custom_theme

B <- S1_summarized_data %>%
  ggplot(aes(
    x = true_prob, y = judged_prob,
    color = new_q_perceived_difficulty_unbounded - observed_q_perceived_difficulty_unbounded
  )) +
  geom_point() +
  scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(
    x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    title = "Participants'\nestimations",
    color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
  ) +
  xlim(0, 1) + ylim(0, 1) +
  custom_theme

C <- S1_summarized_data %>%
  ggplot(aes(
    x = true_prob, y = prob_bayes,
    color = new_q_perceived_difficulty_unbounded - observed_q_perceived_difficulty_unbounded
  )) +
  geom_point() +
  scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(
    x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    title = "Bayesian model's\nestimations",
    color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
  ) +
  xlim(0, 1) + ylim(0, 1) +
  custom_theme

S1_fig_behav <- (A + B + C) +
  plot_layout(guides = "collect", axis = "collect") &
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", axis.title = element_text(size = 9))

S1_fig_behav
```

```{r S1-fig-model-comparison, fig.cap="Model comparison for Study 1. Each data point corresponds to the average prediction for an observed-new question pair. The left column shows participants' average predictions, while the remaining columns show predictions from the Bayesian model, Threshold model, and Anchor model respectively. The top row shows trials where success was observed, and the bottom row shows trials where failure was observed. Lines connect predictions for the same observed question across different evaluated questions, colored by the observed question's perceived difficulty."}
plot_df <- S1_summarized_data %>%
  pivot_longer(
    cols = c(judged_prob, prob_bayes, prob_null, prob_null_2),
    names_to = "model", values_to = "judgment"
  ) %>%
  mutate(
    model = factor(model,
      levels = c("judged_prob", "prob_bayes", "prob_null", "prob_null_2"),
      labels = c("Participant", "Bayesian model", "Threshold model", "Anchor model")
    ),
    success_observed = factor(success_observed,
      levels = c(1, 0),
      labels = c("Success observed", "Failure observed")
    )
  )

S1_fig_model_comparison <- ggplot(
  plot_df,
  aes(
    new_q_perceived_difficulty_unbounded, judgment,
    colour = observed_q_perceived_difficulty_unbounded,
    group  = observed_q_perceived_difficulty_unbounded
  )
) +
  geom_line(alpha = .35, linewidth = .6) +
  geom_point(size = .8) +
  facet_grid(success_observed ~ model) +
  scale_colour_gradientn(
    name    = "Observed Question\nPerceived Difficulty",
    colours = wes_palette("Zissou1", type = "continuous")
  ) +
  labs(x = "New question difficulty", y = "Estimated Probability of Success") +
  custom_theme +
  theme(legend.position = "bottom")

S1_fig_model_comparison
```

```{r S2-load-data}
S2_long_data    <- read_csv(here("study_2", "data", "clean", "S2_long_data.csv"))
S2_prob_df      <- read_csv(here("study_2", "data", "clean", "S2_prob_df.csv"))
S2_long_answers <- read_csv(here("study_2", "data", "clean", "S2_long_answers.csv"))
S2_difficulty   <- read_csv(here("study_2", "data", "clean", "S2_difficulty.csv"))
```

```{r S2-analyses}
# H1: perceived vs. objective difficulty on observed (Quiz 2) items
S2_h1 <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data
)

# H2a / H2b: is time-limit diagnostic in the ground truth?
S2_long_answers$pair <- interaction(
  S2_long_answers$question_observed,
  S2_long_answers$question_evaluated,
  drop = TRUE
)
S2_long_answers$success_q1 <- S2_long_answers$success_evaluated
S2_long_answers$time_observed <- relevel(factor(S2_long_answers$time_observed), ref = "SlowFast")

S2_h2a <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 1)
)
S2_h2b <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 0)
)

# H3: averaged predictions vs. true conditional probabilities
summarized_data <- S2_long_data %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop")

S2_prob_df$time_observed <- as.integer(relevel(factor(S2_prob_df$time_observed), ref = "SlowFast")) - 1

summarized_data <- left_join(
  summarized_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)

S2_h3 <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data)

# H4a / H4b: do participants use the time-limit cue?
S2_long_data$pair <- interaction(
  S2_long_data$question_observed,
  S2_long_data$question_evaluated,
  drop = TRUE
)
S2_h4a <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 1)
)
S2_h4b <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 0)
)

# ---- Robustness (bottom 30%) ----
n_participants <- S2_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S2_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S2_long_data_worst <- S2_long_data %>% filter(prolific_pid %in% worst_ids)

S2_h1m <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data_worst
)

summarized_data_worst <- S2_long_data_worst %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop") %>%
  left_join(S2_prob_df, by = c("question_observed", "success_observed", "time_observed", "question_evaluated"))

S2_h3m <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data_worst)

# Trial-level analysis
S2_long_data <- left_join(
  S2_long_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)
S2_h3i <- glmer(
  human_evaluation ~ scale(true_conditional_prob) +
    (1 | prolific_pid) + (1 | time_observed) + (1 | question_observed) + (1 | question_evaluated),
  family = binomial, S2_long_data
)

S2_results <- list(
  h1  = clean_lmer(S2_h1),
  h2a = clean_lmer(S2_h2a),
  h2b = clean_lmer(S2_h2b),
  h3  = clean_lm(S2_h3),
  h4a = clean_lmer(S2_h4a),
  h4b = clean_lmer(S2_h4b),
  h1m = clean_lmer(S2_h1m),
  h3m = clean_lm(S2_h3m),
  h3i = clean_lmer(S2_h3i)
)
```

# Study 2

**H1 (Difficulty judgments):** $\beta$ = `r S2_results$h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1[["observed_q_objective_difficulty"]]$p.value`.

**H2a (Fast accurate answers are competence cues):** b = `r S2_results$h2a[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2a[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2a[["time_observedFastSlow"]]$p.value`.

**H2b (Fast inaccurate answers are competence cues):** b = `r S2_results$h2b[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2b[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2b[["time_observedFastSlow"]]$p.value`.

**H3 (Competence inferences from a single outcome):** $\beta$ = `r S2_results$h3[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3$model_fit$df.residual`) = `r S2_results$h3[["true_conditional_prob"]]$statistic`, p `r S2_results$h3[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3$model_fit$adj.r.squared`.

**H4a (Participants treat faster accurate answers as competence cues):** b = `r S2_results$h4a[["1"]]$estimate`, z = `r S2_results$h4a[["1"]]$statistic`, p `r S2_results$h4a[["1"]]$p.value`.

**H4b (Participants treat faster inaccurate answers as competence cues):** b = `r S2_results$h4b[["1"]]$estimate`, z = `r S2_results$h4b[["1"]]$statistic`, p `r S2_results$h4b[["1"]]$p.value`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S2_results$h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1m[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S2_results$h3m[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3m$model_fit$df.residual`) = `r S2_results$h3m[["true_conditional_prob"]]$statistic`, p `r S2_results$h3m[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3m$model_fit$adj.r.squared`.
- Trial-level analysis: b = `r S2_results$h3i[["true_conditional_prob"]]$estimate`, z = `r S2_results$h3i[["true_conditional_prob"]]$statistic`, p `r S2_results$h3i[["true_conditional_prob"]]$p.value`.

```{r S2-fig-behav, fig.cap="Estimated Conditional Probability versus True Conditional Probability in Study 2. Each data point is a pair of two questions. The red line represents the optimal prediction while the dark line represents the fit of the linear models. (\\textbf{A}) shows all trials combined, (\\textbf{B}) shows trials in the Fast condition (20-second time limit), and (\\textbf{C}) shows trials in the Slow condition (40-second time limit). Points are colored by the difficulty difference between the evaluated and observed questions.", fig.width = 7, fig.height = 4}
S2_difficulty_obs_fig1 <- S2_difficulty %>%
  select(question, perceived_difficulty) %>%
  rename(observed_q_perceived_difficulty = perceived_difficulty)

# Mapping from S2 Quiz 1 questions to S1 question numbers (based on answer keys)
s2_to_s1_mapping_fig1 <- tibble(
  question_evaluated = c("1_1", "1_2", "1_3", "1_4", "1_5"),
  s1_question = c(2, 6, 8, 13, 15)
)

S1_difficulty_eval_fig1 <- S1_difficulty %>%
  select(question, perceived_difficulty) %>%
  inner_join(s2_to_s1_mapping_fig1, by = c("question" = "s1_question")) %>%
  select(question_evaluated, perceived_difficulty) %>%
  rename(new_q_perceived_difficulty = perceived_difficulty)

S2_summarized_enhanced <- S2_long_data %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarise(
    judg_conditional_prob = mean(human_evaluation, na.rm = TRUE),
    true_conditional_prob = mean(true_conditional_prob, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(S2_difficulty_obs_fig1, by = c("question_observed" = "question")) %>%
  left_join(S1_difficulty_eval_fig1, by = "question_evaluated")

S2_summarized_all  <- S2_summarized_enhanced
S2_summarized_fast <- S2_summarized_enhanced %>% filter(time_observed == 1)
S2_summarized_slow <- S2_summarized_enhanced %>% filter(time_observed == 0)

df_s2_scale <- function(df) {
  df %>%
    mutate(
      x_s = as.numeric(scale(true_conditional_prob)),
      y_s = as.numeric(scale(judg_conditional_prob)),
      diff_col = new_q_perceived_difficulty - observed_q_perceived_difficulty
    )
}

d1_s2 <- df_s2_scale(S2_summarized_all)
d2_s2 <- df_s2_scale(S2_summarized_fast)
d3_s2 <- df_s2_scale(S2_summarized_slow)

p_base_s2 <- function(dat, title_txt) {
  ggplot(dat, aes(x = x_s, y = y_s, color = diff_col)) +
    geom_point() +
    scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    geom_smooth(method = "lm", se = TRUE, color = "black") +
    labs(
      title = title_txt,
      x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
      y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
      color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
    ) +
    custom_theme
}

p1_s2 <- p_base_s2(d1_s2, "A. All trials")
p2_s2 <- p_base_s2(d2_s2, "B. Fast condition (20s)")
p3_s2 <- p_base_s2(d3_s2, "C. Slow condition (40s)")

S2_fig_behav <- p1_s2 + p2_s2 + p3_s2 +
  plot_layout(guides = "collect", axis = "collect") &
  theme(legend.position = "bottom", axis.title = element_text(size = 12))

S2_fig_behav
```
