---
title: "Inline stats for CogSci 2026: 'Inferring Arithmetic Skill from Speed and Accuracy'"
output:
  papaja::apa6_pdf:
    keep_tex: true
  papaja::apa6_docx:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.retina = 3, dpi = 300
)
options(
  width = 90,
  dplyr.summarise.inform = FALSE
)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(papaja)
library(lmerTest)
library(broom)
library(broom.mixed)
library(here)
library(lme4)
```

```{r functions}
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 2)

clean_lm <- function(model_output) {
  result <- tidy(model_output, conf.int = TRUE) %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, statistic, p.value) %>%
    split(.$clean_term)

  model_fit <- glance(model_output) %>%
    select(adj.r.squared, df.residual, df, statistic, p.value) %>%
    rounded_numbers() %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit
  return(result)
}

clean_lmer <- function(model_output) {
  broom.mixed::tidy(model_output, conf.int = TRUE) %>%
    filter(effect == "fixed") %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, any_of("df"), statistic, p.value) %>%
    split(.$clean_term)
}
```

```{r S1-load-data}
S1_long_data <- read_csv(here::here("study_1", "data", "clean", "S1_long_data.csv"))
S1_prob_df   <- read_csv(here::here("study_1", "data", "clean", "S1_prob_df.csv"))

source(here::here("pre-registered code", "pre-reg_model_implementation.R"))

S1_fit_results <- read_csv(here::here("study_1", "results", "S1_fit_results.csv"))
```

```{r S1-analyses}
S1_long_data <- left_join(
  S1_long_data, S1_prob_df,
  by = c("question_observed", "success_observed", "question_evaluated")
)

# Bottom 30% subsample (robustness)
n_participants <- S1_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S1_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S1_long_data_worst <- S1_long_data %>% filter(prolific_pid %in% worst_ids)

# Summarized data (one row per observed-new question pair x success/failure)
S1_summarized_data <- S1_long_data %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    judged_prob = mean(human_evaluation, na.rm = TRUE),
    observed_q_perceived_difficulty_unbounded = mean(observed_q_perceived_difficulty_unbounded),
    new_q_perceived_difficulty_unbounded = mean(new_q_perceived_difficulty_unbounded),
    true_prob = mean(true_conditional_prob),
    .groups = "drop"
  ) %>%
  filter(question_observed != question_evaluated)

# ---- Bayesian / heuristic model predictions ----
S1_MAIN_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$LLMax + 4 * log(nrow(S1_long_data))
S1_NULL_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL", ]$LLMax + 1 * log(nrow(S1_long_data))
S1_NULL_2_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$LLMax + 2 * log(nrow(S1_long_data))

delta <- 0.005
COMP  <- seq(-4, 4, by = delta)

mu    <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$mu
sigma <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$sigma
noise <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$noise
beta  <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$beta

S1_prior_COMP <- dnorm(COMP, mu, sigma)
S1_prior_COMP <- S1_prior_COMP / sum(S1_prior_COMP * delta)

S1_summarized_data$prob_bayes <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_prior_COMP, noise = noise, beta = beta)
)

beta <- S1_fit_results[S1_fit_results$model == "S1_NULL", ]$beta
S1_NULL_COMP <- rep(1, length(COMP))
S1_NULL_COMP <- S1_NULL_COMP / sum(S1_NULL_COMP * delta)

S1_summarized_data$prob_null <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_NULL_COMP, noise = 0, beta = beta)
)

beta       <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$beta
difference <- beta <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$diff

S1_summarized_data$prob_null_2 <- mapply(
  FUN = function(obs, new, success_observed) {
    if (success_observed == 1) {
      soft_max(ifelse(obs + difference >= new, 1, 0), beta = beta)
    } else {
      soft_max(ifelse(obs - difference >= new, 1, 0), beta = beta)
    }
  },
  obs = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success_observed = S1_summarized_data$success_observed
)

# ---- Confirmatory models ----
data_h1 <- S1_long_data %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty_indiv) %>%
  distinct()
model_h1 <- lmer(scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1)

model_h2 <- lm(scale(judged_prob) ~ scale(true_prob), data = S1_summarized_data)

model_bayes <- lm(scale(judged_prob) ~ scale(prob_bayes), data = S1_summarized_data)
model_null1 <- lm(scale(judged_prob) ~ scale(prob_null),  data = S1_summarized_data)
model_null2 <- lm(scale(judged_prob) ~ scale(prob_null_2), data = S1_summarized_data)

# ---- Robustness (bottom 30%) ----
data_h1_worst <- S1_long_data_worst %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty) %>%
  distinct()
model_h1m <- lmer(scale(observed_q_perceived_difficulty) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1_worst)

data_h2m <- S1_long_data_worst %>%
  group_by(question_observed, question_evaluated, success_observed) %>%
  summarise(
    true_prob   = mean(true_conditional_prob),
    judged_prob = mean(human_evaluation),
    .groups = "drop"
  )
model_h2m <- lm(scale(judged_prob) ~ scale(true_prob), data_h2m)

S1_results <- list(
  clean_model_h1    = clean_lmer(model_h1),
  clean_model_h2    = clean_lm(model_h2),
  clean_model_bayes = clean_lm(model_bayes),
  clean_model_null1 = clean_lm(model_null1),
  clean_model_null2 = clean_lm(model_null2),
  clean_model_h1m   = clean_lmer(model_h1m),
  clean_model_h2m   = clean_lm(model_h2m)
)
```

# Study 1

**H1 (Difficulty judgments):** $\beta$ = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$p.value`.

**H2 (Competence inferences from a single outcome):** $\beta$ = `r S1_results$clean_model_h2$true_prob$estimate`, t(`r S1_results$clean_model_h2$model_fit$df.residual`) = `r S1_results$clean_model_h2$true_prob$statistic`, p `r S1_results$clean_model_h2$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2$model_fit$adj.r.squared`.

**H3--H4 (Model accuracy and comparison):**

- Main (Bayesian) Model: $\beta$ = `r S1_results$clean_model_bayes$prob_bayes$estimate`, t(`r S1_results$clean_model_bayes$model_fit$df.residual`) = `r S1_results$clean_model_bayes$prob_bayes$statistic`, p `r S1_results$clean_model_bayes$prob_bayes$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_bayes$model_fit$adj.r.squared`.
- Threshold: Adjusted $R^2$ = `r S1_results$clean_model_null1$model_fit$adj.r.squared`.
- Anchor: Adjusted $R^2$ = `r S1_results$clean_model_null2$model_fit$adj.r.squared`.
- BIC: $BIC_{\text{Bayes}}$ = `r round(S1_MAIN_BIC, 2)`; $BIC_{\text{Threshold}}$ = `r round(S1_NULL_BIC, 2)`; $BIC_{\text{Anchor}}$ = `r round(S1_NULL_2_BIC, 2)`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S1_results$clean_model_h2m$true_prob$estimate`, t(`r S1_results$clean_model_h2m$model_fit$df.residual`) = `r S1_results$clean_model_h2m$true_prob$statistic`, p `r S1_results$clean_model_h2m$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2m$model_fit$adj.r.squared`.

```{r S2-load-data}
S2_long_data    <- read_csv(here("study_2", "data", "clean", "S2_long_data.csv"))
S2_prob_df      <- read_csv(here("study_2", "data", "clean", "S2_prob_df.csv"))
S2_long_answers <- read_csv(here("study_2", "data", "clean", "S2_long_answers.csv"))
```

```{r S2-analyses}
# H1: perceived vs. objective difficulty on observed (Quiz 2) items
S2_h1 <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data
)

# H2a / H2b: is time-limit diagnostic in the ground truth?
S2_long_answers$pair <- interaction(
  S2_long_answers$question_observed,
  S2_long_answers$question_evaluated,
  drop = TRUE
)
S2_long_answers$success_q1 <- S2_long_answers$success_evaluated
S2_long_answers$time_observed <- relevel(factor(S2_long_answers$time_observed), ref = "SlowFast")

S2_h2a <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 1)
)
S2_h2b <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 0)
)

# H3: averaged predictions vs. true conditional probabilities
summarized_data <- S2_long_data %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop")

S2_prob_df$time_observed <- as.integer(relevel(factor(S2_prob_df$time_observed), ref = "SlowFast")) - 1

summarized_data <- left_join(
  summarized_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)

S2_h3 <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data)

# H4a / H4b: do participants use the time-limit cue?
S2_long_data$pair <- interaction(
  S2_long_data$question_observed,
  S2_long_data$question_evaluated,
  drop = TRUE
)
S2_h4a <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 1)
)
S2_h4b <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 0)
)

# ---- Robustness (bottom 30%) ----
n_participants <- S2_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S2_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S2_long_data_worst <- S2_long_data %>% filter(prolific_pid %in% worst_ids)

S2_h1m <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data_worst
)

summarized_data_worst <- S2_long_data_worst %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop") %>%
  left_join(S2_prob_df, by = c("question_observed", "success_observed", "time_observed", "question_evaluated"))

S2_h3m <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data_worst)

# Trial-level analysis
S2_long_data <- left_join(
  S2_long_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)
S2_h3i <- glmer(
  human_evaluation ~ scale(true_conditional_prob) +
    (1 | prolific_pid) + (1 | time_observed) + (1 | question_observed) + (1 | time_observed),
  family = binomial, S2_long_data
)

S2_results <- list(
  h1  = clean_lmer(S2_h1),
  h2a = clean_lmer(S2_h2a),
  h2b = clean_lmer(S2_h2b),
  h3  = clean_lm(S2_h3),
  h4a = clean_lmer(S2_h4a),
  h4b = clean_lmer(S2_h4b),
  h1m = clean_lmer(S2_h1m),
  h3m = clean_lm(S2_h3m),
  h3i = clean_lmer(S2_h3i)
)
```

# Study 2

**H1 (Difficulty judgments):** $\beta$ = `r S2_results$h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1[["observed_q_objective_difficulty"]]$p.value`.

**H2a (Fast accurate answers are competence cues):** b = `r S2_results$h2a[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2a[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2a[["time_observedFastSlow"]]$p.value`.

**H2b (Fast inaccurate answers are competence cues):** b = `r S2_results$h2b[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2b[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2b[["time_observedFastSlow"]]$p.value`.

**H3 (Competence inferences from a single outcome):** $\beta$ = `r S2_results$h3[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3$model_fit$df.residual`) = `r S2_results$h3[["true_conditional_prob"]]$statistic`, p `r S2_results$h3[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3$model_fit$adj.r.squared`.

**H4a (Participants treat faster accurate answers as competence cues):** b = `r S2_results$h4a[["1"]]$estimate`, z = `r S2_results$h4a[["1"]]$statistic`, p `r S2_results$h4a[["1"]]$p.value`.

**H4b (Participants treat faster inaccurate answers as competence cues):** b = `r S2_results$h4b[["1"]]$estimate`, z = `r S2_results$h4b[["1"]]$statistic`, p `r S2_results$h4b[["1"]]$p.value`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S2_results$h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1m[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S2_results$h3m[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3m$model_fit$df.residual`) = `r S2_results$h3m[["true_conditional_prob"]]$statistic`, p `r S2_results$h3m[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3m$model_fit$adj.r.squared`.
- Trial-level analysis: b = `r S2_results$h3i[["true_conditional_prob"]]$estimate`, z = `r S2_results$h3i[["true_conditional_prob"]]$statistic`, p `r S2_results$h3i[["true_conditional_prob"]]$p.value`.
