---
title: "Analysis code for CogSci 2026: 'Inferring Arithmetic Skill from Speed and Accuracy'"
output:
  papaja::apa6_pdf:
    keep_tex: true
  papaja::apa6_docx:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.retina = 3, dpi = 300
)
options(
  width = 90,
  dplyr.summarise.inform = FALSE
)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(papaja)
library(lmerTest)
library(broom)
library(broom.mixed)
library(here)
library(lme4)
library(patchwork)
library(ggrepel)
library(wesanderson)
```

```{r theme}
custom_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    strip.text = element_text(colour = "black"),
    strip.background   = element_rect(fill = "white", colour = "black", linewidth = .8),
    strip.background.x = element_rect(fill = "white", colour = "black", linewidth = .8),
    strip.background.y = element_rect(fill = "white", colour = "black", linewidth = .8)
  )
```

```{r functions}
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 2)

clean_lm <- function(model_output) {
  result <- tidy(model_output, conf.int = TRUE) %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, statistic, p.value) %>%
    split(.$clean_term)

  model_fit <- glance(model_output) %>%
    select(adj.r.squared, df.residual, df, statistic, p.value) %>%
    rounded_numbers() %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit
  return(result)
}

clean_lmer <- function(model_output) {
  broom.mixed::tidy(model_output, conf.int = TRUE) %>%
    filter(effect == "fixed") %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, any_of("df"), statistic, p.value) %>%
    split(.$clean_term)
}
```

```{r S1-load-data}
S1_long_data   <- read_csv(here::here("study_1", "data", "clean", "S1_long_data.csv"))
S1_prob_df     <- read_csv(here::here("study_1", "data", "clean", "S1_prob_df.csv"))
S1_difficulty  <- read_csv(here::here("study_1", "data", "clean", "S1_difficulty.csv"))

source(here::here("pre-registered code", "pre-reg_model_implementation.R"))

S1_fit_results <- read_csv(here::here("study_1", "results", "S1_fit_results.csv"))
```

```{r S1-analyses}
S1_long_data <- left_join(
  S1_long_data, S1_prob_df,
  by = c("question_observed", "success_observed", "question_evaluated")
)

# Bottom 30% subsample (robustness)
n_participants <- S1_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S1_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S1_long_data_worst <- S1_long_data %>% filter(prolific_pid %in% worst_ids)

# Summarized data (one row per observed-new question pair x success/failure)
S1_summarized_data <- S1_long_data %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    judged_prob = mean(human_evaluation, na.rm = TRUE),
    observed_q_perceived_difficulty_unbounded = mean(observed_q_perceived_difficulty_unbounded),
    new_q_perceived_difficulty_unbounded = mean(new_q_perceived_difficulty_unbounded),
    true_prob = mean(true_conditional_prob),
    .groups = "drop"
  ) %>%
  filter(question_observed != question_evaluated)

# ---- Bayesian / heuristic model predictions ----
S1_MAIN_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$LLMax + 4 * log(nrow(S1_long_data))
S1_NULL_BIC   <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL", ]$LLMax + 1 * log(nrow(S1_long_data))
S1_NULL_2_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$LLMax + 2 * log(nrow(S1_long_data))

delta <- 0.005
COMP  <- seq(-4, 4, by = delta)

mu    <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$mu
sigma <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$sigma
noise <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$noise
beta  <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$beta

S1_prior_COMP <- dnorm(COMP, mu, sigma)
S1_prior_COMP <- S1_prior_COMP / sum(S1_prior_COMP * delta)

S1_summarized_data$prob_bayes <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_prior_COMP, noise = noise, beta = beta)
)

beta <- S1_fit_results[S1_fit_results$model == "S1_NULL", ]$beta
S1_NULL_COMP <- rep(1, length(COMP))
S1_NULL_COMP <- S1_NULL_COMP / sum(S1_NULL_COMP * delta)

S1_summarized_data$prob_null <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded      = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success = S1_summarized_data$success_observed,
  MoreArgs = list(prior = S1_NULL_COMP, noise = 0, beta = beta)
)

beta       <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$beta
difference <- beta <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$diff

S1_summarized_data$prob_null_2 <- mapply(
  FUN = function(obs, new, success_observed) {
    if (success_observed == 1) {
      soft_max(ifelse(obs + difference >= new, 1, 0), beta = beta)
    } else {
      soft_max(ifelse(obs - difference >= new, 1, 0), beta = beta)
    }
  },
  obs = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  success_observed = S1_summarized_data$success_observed
)

# ---- Confirmatory models ----
data_h1 <- S1_long_data %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty_indiv) %>%
  distinct()
model_h1 <- lmer(scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1)

model_h2 <- lm(scale(judged_prob) ~ scale(true_prob), data = S1_summarized_data)

model_bayes <- lm(scale(judged_prob) ~ scale(prob_bayes), data = S1_summarized_data)
model_null1 <- lm(scale(judged_prob) ~ scale(prob_null),  data = S1_summarized_data)
model_null2 <- lm(scale(judged_prob) ~ scale(prob_null_2), data = S1_summarized_data)

# ---- Robustness (bottom 30%) ----
data_h1_worst <- S1_long_data_worst %>%
  select(prolific_pid, question_observed, observed_q_objective_difficulty, observed_q_perceived_difficulty) %>%
  distinct()
model_h1m <- lmer(scale(observed_q_perceived_difficulty) ~ scale(observed_q_objective_difficulty) + (1 | prolific_pid), data_h1_worst)

data_h2m <- S1_long_data_worst %>%
  group_by(question_observed, question_evaluated, success_observed) %>%
  summarise(
    true_prob   = mean(true_conditional_prob),
    judged_prob = mean(human_evaluation),
    .groups = "drop"
  )
model_h2m <- lm(scale(judged_prob) ~ scale(true_prob), data_h2m)

S1_results <- list(
  clean_model_h1    = clean_lmer(model_h1),
  clean_model_h2    = clean_lm(model_h2),
  clean_model_bayes = clean_lm(model_bayes),
  clean_model_null1 = clean_lm(model_null1),
  clean_model_null2 = clean_lm(model_null2),
  clean_model_h1m   = clean_lmer(model_h1m),
  clean_model_h2m   = clean_lm(model_h2m)
)
```

# Study 1

**H1 (Difficulty judgments):** $\beta$ = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1[["observed_q_objective_difficulty"]]$p.value`.

**H2 (Competence inferences from a single outcome):** $\beta$ = `r S1_results$clean_model_h2$true_prob$estimate`, t(`r S1_results$clean_model_h2$model_fit$df.residual`) = `r S1_results$clean_model_h2$true_prob$statistic`, p `r S1_results$clean_model_h2$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2$model_fit$adj.r.squared`.

**H3--H4 (Model accuracy and comparison):**

- Main (Bayesian) Model: $\beta$ = `r S1_results$clean_model_bayes$prob_bayes$estimate`, t(`r S1_results$clean_model_bayes$model_fit$df.residual`) = `r S1_results$clean_model_bayes$prob_bayes$statistic`, p `r S1_results$clean_model_bayes$prob_bayes$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_bayes$model_fit$adj.r.squared`.
- Threshold: Adjusted $R^2$ = `r S1_results$clean_model_null1$model_fit$adj.r.squared`.
- Anchor: Adjusted $R^2$ = `r S1_results$clean_model_null2$model_fit$adj.r.squared`.
- BIC: $BIC_{\text{Bayes}}$ = `r round(S1_MAIN_BIC, 2)`; $BIC_{\text{Threshold}}$ = `r round(S1_NULL_BIC, 2)`; $BIC_{\text{Anchor}}$ = `r round(S1_NULL_2_BIC, 2)`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$df`) = `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S1_results$clean_model_h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S1_results$clean_model_h2m$true_prob$estimate`, t(`r S1_results$clean_model_h2m$model_fit$df.residual`) = `r S1_results$clean_model_h2m$true_prob$statistic`, p `r S1_results$clean_model_h2m$true_prob$p.value`, Adjusted $R^2$ = `r S1_results$clean_model_h2m$model_fit$adj.r.squared`.

```{r S1-fig-behav, fig.cap = "(\\textbf{A}) Each question's estimated difficulty as a function of its true difficulty. Numbers refer to the question ID (see Materials). The red dashed line represents a perfectly accurate estimation. (\\textbf{B}) Participants' estimated probability of answering a new question correctly, conditioned on the performance on the observed question, as a function of true conditional probabilities. (\\textbf{C}) The Bayesian model's estimated conditional probabilities as a function of true conditional probabilities. For (\\textbf{B}) and (\\textbf{C}), each data point corresponds to the average prediction for an observed-new question pair. The red line represents a perfectly accurate prediction, while the dark line represents the fit of the linear model.", fig.width = 7, fig.height = 4}
A <- S1_difficulty %>%
  ggplot(aes(x = objective_difficulty, y = perceived_difficulty)) +
  geom_point(color = "#E4A70A") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  geom_label_repel(
    aes(label = question),
    color = "#E4A70A",
    size = 2.5, show.legend = FALSE, seed = 42,
    max.overlaps = Inf, box.padding = 0.4, point.padding = 0.15,
    label.padding = grid::unit(0.12, "lines"),
    label.size = 0.3, label.r = grid::unit(0.15, "lines"),
    fill = scales::alpha("white", 0.7),
    force = 3, force_pull = 0.2, max.iter = 50000,
    min.segment.length = 0, segment.color = "grey60", segment.size = 0.3
  ) +
  xlim(0, 1) + ylim(0, 1) +
  labs(x = "True difficulty", y = "Estimated difficulty", title = "Participants'\nestimations") +
  custom_theme

B <- S1_summarized_data %>%
  ggplot(aes(
    x = true_prob, y = judged_prob,
    color = new_q_perceived_difficulty_unbounded - observed_q_perceived_difficulty_unbounded
  )) +
  geom_point() +
  scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(
    x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    title = "Participants'\nestimations",
    color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
  ) +
  xlim(0, 1) + ylim(0, 1) +
  custom_theme

C <- S1_summarized_data %>%
  ggplot(aes(
    x = true_prob, y = prob_bayes,
    color = new_q_perceived_difficulty_unbounded - observed_q_perceived_difficulty_unbounded
  )) +
  geom_point() +
  scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(
    x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
    title = "Bayesian model's\nestimations",
    color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
  ) +
  xlim(0, 1) + ylim(0, 1) +
  custom_theme

S1_fig_behav <- (A + B + C) +
  plot_layout(guides = "collect", axis = "collect") &
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", axis.title = element_text(size = 9))

S1_fig_behav
```

```{r S1-fig-model-comparison, fig.cap="Model comparison for Study 1. Each data point corresponds to the average prediction for an observed-new question pair. The left column shows participants' average predictions, while the remaining columns show predictions from the Bayesian model, Threshold model, and Anchor model respectively. The top row shows trials where success was observed, and the bottom row shows trials where failure was observed. Lines connect predictions for the same observed question across different evaluated questions, colored by the observed question's perceived difficulty."}
plot_df <- S1_summarized_data %>%
  pivot_longer(
    cols = c(judged_prob, prob_bayes, prob_null, prob_null_2),
    names_to = "model", values_to = "judgment"
  ) %>%
  mutate(
    model = factor(model,
      levels = c("judged_prob", "prob_bayes", "prob_null", "prob_null_2"),
      labels = c("Participant", "Bayesian model", "Threshold model", "Anchor model")
    ),
    success_observed = factor(success_observed,
      levels = c(1, 0),
      labels = c("Success observed", "Failure observed")
    )
  )

S1_fig_model_comparison <- ggplot(
  plot_df,
  aes(
    new_q_perceived_difficulty_unbounded, judgment,
    colour = observed_q_perceived_difficulty_unbounded,
    group  = observed_q_perceived_difficulty_unbounded
  )
) +
  geom_line(alpha = .35, linewidth = .6) +
  geom_point(size = .8) +
  facet_grid(success_observed ~ model) +
  scale_colour_gradientn(
    name    = "Observed Question\nPerceived Difficulty",
    colours = wes_palette("Zissou1", type = "continuous")
  ) +
  labs(x = "New question difficulty", y = "Estimated Probability of Success") +
  custom_theme +
  theme(legend.position = "bottom")

S1_fig_model_comparison
```

```{r S2-load-data}
S2_long_data    <- read_csv(here("study_2", "data", "clean", "S2_long_data.csv"))
S2_prob_df      <- read_csv(here("study_2", "data", "clean", "S2_prob_df.csv"))
S2_long_answers <- read_csv(here("study_2", "data", "clean", "S2_long_answers.csv"))
S2_difficulty   <- read_csv(here("study_2", "data", "clean", "S2_difficulty.csv"))
```

```{r S2-analyses}
# H1: perceived vs. objective difficulty on observed (Quiz 2) items
S2_h1 <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data
)

# H2a / H2b: is time-limit diagnostic in the ground truth?
S2_long_answers$pair <- interaction(
  S2_long_answers$question_observed,
  S2_long_answers$question_evaluated,
  drop = TRUE
)
S2_long_answers$success_q1 <- S2_long_answers$success_evaluated
S2_long_answers$time_observed <- relevel(factor(S2_long_answers$time_observed), ref = "SlowFast")

S2_h2a <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 1)
)
S2_h2b <- glmer(
  success_q1 ~ time_observed + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_answers %>% filter(success_observed == 0)
)

# H3: averaged predictions vs. true conditional probabilities
summarized_data <- S2_long_data %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop")

S2_prob_df$time_observed <- as.integer(relevel(factor(S2_prob_df$time_observed), ref = "SlowFast")) - 1

summarized_data <- left_join(
  summarized_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)

S2_h3 <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data)

# H4a / H4b: do participants use the time-limit cue?
S2_long_data$pair <- interaction(
  S2_long_data$question_observed,
  S2_long_data$question_evaluated,
  drop = TRUE
)
S2_h4a <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 1)
)
S2_h4b <- glmer(
  human_evaluation ~ factor(time_observed) + (1 | prolific_pid) + (1 | pair),
  family = binomial, S2_long_data %>% filter(success_observed == 0)
)

# ---- Robustness (bottom 30%) ----
n_participants <- S2_long_data %>% distinct(prolific_pid) %>% nrow()
n_worst <- floor(n_participants * 0.30)
worst_ids <- S2_long_data %>%
  distinct(prolific_pid, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(prolific_pid)
S2_long_data_worst <- S2_long_data %>% filter(prolific_pid %in% worst_ids)

S2_h1m <- lmer(
  scale(observed_q_perceived_difficulty_indiv) ~ scale(observed_q_objective_difficulty) +
    (1 + scale(observed_q_objective_difficulty) | prolific_pid),
  S2_long_data_worst
)

summarized_data_worst <- S2_long_data_worst %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarize(judg_conditional_prob = mean(human_evaluation, na.rm = TRUE), .groups = "drop") %>%
  left_join(S2_prob_df, by = c("question_observed", "success_observed", "time_observed", "question_evaluated"))

S2_h3m <- lm(scale(judg_conditional_prob) ~ scale(true_conditional_prob), data = summarized_data_worst)

# Trial-level analysis
S2_long_data <- left_join(
  S2_long_data, S2_prob_df,
  by = c("question_observed", "success_observed", "time_observed", "question_evaluated")
)
S2_h3i <- glmer(
  human_evaluation ~ scale(true_conditional_prob) +
    (1 | prolific_pid) + (1 | time_observed) + (1 | question_observed) + (1 | time_observed),
  family = binomial, S2_long_data
)

S2_results <- list(
  h1  = clean_lmer(S2_h1),
  h2a = clean_lmer(S2_h2a),
  h2b = clean_lmer(S2_h2b),
  h3  = clean_lm(S2_h3),
  h4a = clean_lmer(S2_h4a),
  h4b = clean_lmer(S2_h4b),
  h1m = clean_lmer(S2_h1m),
  h3m = clean_lm(S2_h3m),
  h3i = clean_lmer(S2_h3i)
)
```

# Study 2

**H1 (Difficulty judgments):** $\beta$ = `r S2_results$h1[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1[["observed_q_objective_difficulty"]]$p.value`.

**H2a (Fast accurate answers are competence cues):** b = `r S2_results$h2a[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2a[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2a[["time_observedFastSlow"]]$p.value`.

**H2b (Fast inaccurate answers are competence cues):** b = `r S2_results$h2b[["time_observedFastSlow"]]$estimate`, z = `r S2_results$h2b[["time_observedFastSlow"]]$statistic`, p `r S2_results$h2b[["time_observedFastSlow"]]$p.value`.

**H3 (Competence inferences from a single outcome):** $\beta$ = `r S2_results$h3[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3$model_fit$df.residual`) = `r S2_results$h3[["true_conditional_prob"]]$statistic`, p `r S2_results$h3[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3$model_fit$adj.r.squared`.

**H4a (Participants treat faster accurate answers as competence cues):** b = `r S2_results$h4a[["1"]]$estimate`, z = `r S2_results$h4a[["1"]]$statistic`, p `r S2_results$h4a[["1"]]$p.value`.

**H4b (Participants treat faster inaccurate answers as competence cues):** b = `r S2_results$h4b[["1"]]$estimate`, z = `r S2_results$h4b[["1"]]$statistic`, p `r S2_results$h4b[["1"]]$p.value`.

**Robustness (bottom 30% of arithmetic performers):**

- Perceived vs. objective difficulty: $\beta$ = `r S2_results$h1m[["observed_q_objective_difficulty"]]$estimate`, t(`r S2_results$h1m[["observed_q_objective_difficulty"]]$df`) = `r S2_results$h1m[["observed_q_objective_difficulty"]]$statistic`, p `r S2_results$h1m[["observed_q_objective_difficulty"]]$p.value`.
- Conditional predictions: $\beta$ = `r S2_results$h3m[["true_conditional_prob"]]$estimate`, t(`r S2_results$h3m$model_fit$df.residual`) = `r S2_results$h3m[["true_conditional_prob"]]$statistic`, p `r S2_results$h3m[["true_conditional_prob"]]$p.value`, Adjusted $R^2$ = `r S2_results$h3m$model_fit$adj.r.squared`.
- Trial-level analysis: b = `r S2_results$h3i[["true_conditional_prob"]]$estimate`, z = `r S2_results$h3i[["true_conditional_prob"]]$statistic`, p `r S2_results$h3i[["true_conditional_prob"]]$p.value`.

```{r S2-fig-behav, fig.cap="Estimated Conditional Probability versus True Conditional Probability in Study 2. Each data point is a pair of two questions. The red line represents the optimal prediction while the dark line represents the fit of the linear models. (\\textbf{A}) shows all trials combined, (\\textbf{B}) shows trials in the Fast condition (20-second time limit), and (\\textbf{C}) shows trials in the Slow condition (40-second time limit). Points are colored by the difficulty difference between the evaluated and observed questions.", fig.width = 7, fig.height = 4}
S2_difficulty_obs_fig1 <- S2_difficulty %>%
  select(question, perceived_difficulty) %>%
  rename(observed_q_perceived_difficulty = perceived_difficulty)

# Mapping from S2 Quiz 1 questions to S1 question numbers (based on answer keys)
s2_to_s1_mapping_fig1 <- tibble(
  question_evaluated = c("1_1", "1_2", "1_3", "1_4", "1_5"),
  s1_question = c(2, 6, 8, 13, 15)
)

S1_difficulty_eval_fig1 <- S1_difficulty %>%
  select(question, perceived_difficulty) %>%
  inner_join(s2_to_s1_mapping_fig1, by = c("question" = "s1_question")) %>%
  select(question_evaluated, perceived_difficulty) %>%
  rename(new_q_perceived_difficulty = perceived_difficulty)

S2_summarized_enhanced <- S2_long_data %>%
  group_by(question_observed, success_observed, time_observed, question_evaluated) %>%
  summarise(
    judg_conditional_prob = mean(human_evaluation, na.rm = TRUE),
    true_conditional_prob = mean(true_conditional_prob, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(S2_difficulty_obs_fig1, by = c("question_observed" = "question")) %>%
  left_join(S1_difficulty_eval_fig1, by = "question_evaluated")

S2_summarized_all  <- S2_summarized_enhanced
S2_summarized_fast <- S2_summarized_enhanced %>% filter(time_observed == 1)
S2_summarized_slow <- S2_summarized_enhanced %>% filter(time_observed == 0)

df_s2_scale <- function(df) {
  df %>%
    mutate(
      x_s = as.numeric(scale(true_conditional_prob)),
      y_s = as.numeric(scale(judg_conditional_prob)),
      diff_col = new_q_perceived_difficulty - observed_q_perceived_difficulty
    )
}

d1_s2 <- df_s2_scale(S2_summarized_all)
d2_s2 <- df_s2_scale(S2_summarized_fast)
d3_s2 <- df_s2_scale(S2_summarized_slow)

p_base_s2 <- function(dat, title_txt) {
  ggplot(dat, aes(x = x_s, y = y_s, color = diff_col)) +
    geom_point() +
    scale_color_gradient2(low = "blue", mid = "grey", high = "red") +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    geom_smooth(method = "lm", se = TRUE, color = "black") +
    labs(
      title = title_txt,
      x = expression("True " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
      y = expression("Estimated " * italic(P) * "(" * italic(S)[new] == 1 ~ "|" ~ italic(S)[obs] * ", " * italic(beta)[obs] * ", " * italic(beta)[new] * ")"),
      color = expression(atop(atop("Difficulty", "difference"), "(" * italic(beta)[new] ~ "-" ~ italic(beta)[obs] * ")"))
    ) +
    custom_theme
}

p1_s2 <- p_base_s2(d1_s2, "A. All trials")
p2_s2 <- p_base_s2(d2_s2, "B. Fast condition (20s)")
p3_s2 <- p_base_s2(d3_s2, "C. Slow condition (40s)")

S2_fig_behav <- p1_s2 + p2_s2 + p3_s2 +
  plot_layout(guides = "collect", axis = "collect") &
  theme(legend.position = "bottom", axis.title = element_text(size = 12))

S2_fig_behav
```
