---
title: "Result reports of 'Inferring Competence on a Numerical Task'"
author: 
  - name: ""
    affiliation: ""
affiliation:
    
  - id: ""
    institution: ""
date: "`r Sys.Date()`"
header-includes:  |
  \usepackage{placeins}
floatsintext: yes
linenumbers: no 
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
classoption: "doc"
output: 
  # papaja::apa6_pdf:
  #   keep_tex: true
  papaja::apa6_docx: default
always_allow_html: true
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "figures/"
  # cache.path = "output/_cache/",
  # fig.process = function(x) {  # Remove "-1" from figure names
  #   x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  #   if (file.rename(x, x2)) x2 else x
  # },
  # options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)

```

```{r libraries, include=FALSE}
library(tidyverse)
library(mlogit)       # For multinomial logistic regression analysis
library(flextable)    # For creating formatted tables
library(papaja)       # For APA-style formatting
library(lmerTest)     # For linear mixed-effects models with p-values
library(broom)        # For tidying model outputs
library(broom.mixed)  # For tidying mixed effects model outputs
library(wesanderson)  # For color palettes inspired by Wes Anderson films
library(patchwork)    # For combining multiple ggplots
library(here)         # For project-relative file paths
library(ggh4x)        # For extended ggplot2 functionality


# library(gtsummary)
# library(table1)
# library(ggdist)
# library(jtools)
```

```{r functions}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 2)

# get model outputs ready for inline text reporting
clean_lm <- function(model_output) {
  result <- tidy(model_output, conf.int = TRUE) %>%
    # filter(effect == "fixed") %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      # if there is an interaction (recognized by ":"), name it just interaction
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, statistic, p.value) %>%
    split(.$clean_term)

  model_fit <- glance(model_output) %>%
    select(adj.r.squared, df.residual, df, statistic, p.value) %>%
    rounded_numbers() %>%
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit

  return(result)
}

clean_lmer <- function(model_output) {
  result <- broom.mixed::tidy(model_output, conf.int = TRUE) %>%
    filter(effect == "fixed") %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ paste0("= ", sprintf("%.3f", p.value))
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate(
      term = ifelse(term == "(Intercept)", "intercept", term),
      ci = glue::glue("[{conf.low}, {conf.high}]"),
      # if there is an interaction (recognized by ":"), name it just interaction
      term = str_replace(term, ".*:.*", "interaction"),
      clean_term = term %>%
        str_replace_all("factor\\(.*?\\)", "") %>%
        str_replace_all("scale\\((.*?)\\)", "\\1")
    ) %>%
    select(clean_term, term, estimate, std.error, ci, df, statistic, p.value) %>%
    split(.$clean_term)

  return(result)
}

# clean data frame results for paired t-test
clean_t_test <- function(t.test) {
  t.test %>%
    tidy() %>%
    select(-c(method, alternative)) %>%
    # report p.value according to apa standards
    mutate(p.value = case_when(
      p.value < 0.001 ~ "< .001",
      TRUE ~ sprintf("p = %.3f", p.value)
    )) %>%
    # all other terms
    rounded_numbers() %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })
}

clean_mlogit <- function(model_output) {
  # Get the summary of the mlogit model, which includes a coefficient table
  summ <- summary(model_output)

  # Convert the coefficient matrix to a data frame with row names as a column
  coef_table <- as.data.frame(summ$CoefTable)
  coef_table <- tibble::rownames_to_column(coef_table, var = "term")

  # Rename columns to common names (make sure these match your summary output)
  names(coef_table) <- c("term", "estimate", "std.error", "z.value", "p.value")

  # Process the coefficient table
  coef_table <- coef_table %>%
    mutate(
      # Format p-values per APA guidelines
      p.value = case_when(
        p.value < 0.001 ~ "< .001",
        TRUE ~ paste0("= ", sprintf("%.3f", p.value))
      ),
      # Calculate confidence intervals (normal approximation)
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      ci = glue::glue("[{round(conf.low, 2)}, {round(conf.high, 2)}]"),
      # Rename intercept and collapse any interaction terms
      term = ifelse(term == "(Intercept)", "intercept", term),
      term = ifelse(str_detect(term, ":"), "interaction", term)
    ) %>%
    select(term, estimate, std.error, ci, z.value, p.value) %>%
    rounded_numbers()

  # Split the table by term so each coefficient gets its own list element
  result <- split(coef_table, coef_table$term)

  # Create a model fit summary.
  # Here we include logLik, AIC, BIC, and the number of iterations (if available)
  model_fit <- data.frame(
    logLik = as.numeric(model_output$logLik),
    AIC = AIC(model_output),
    BIC = -2 * as.numeric(model_output$logLik) + attr(logLik(model_output), "df") * log(length(model_output$fitted.values)),
    nb.iter = model_output$est.stat$nb.iter
  )
  model_fit <- rounded_numbers(model_fit) %>%
    mutate_all(~ {
      attributes(.) <- NULL
      .
    })

  result$model_fit <- model_fit

  return(result)
}

custom_theme <- theme_minimal() +
  theme(text = element_text(family = "serif"), 
        ## ---- facet-label styling ----
      strip.text        = element_text(colour = "black"),         # text colour
      strip.background  = element_rect(fill = "white",            # inside of the strip
                                      colour = "black",          # border line
                                      linewidth = .8),

      ## ggplot2 ≥ 3.4 splits strips into x- and y-orientations;
      ## setting these keeps the look identical for both directions
      strip.background.x = element_rect(fill = "white", colour = "black", linewidth = .8),
      strip.background.y = element_rect(fill = "white", colour = "black", linewidth = .8)
  )
        # panel.background = element_rect(fill = "white", colour = NA), 
        # panel.border     = element_rect(fill = NA, colour = "black", linewidth = .4)) 

# Theme for plots in slides:
custom_theme_dark <- theme_minimal() +
  theme(
    text = element_text(family = "serif", color = "white", size = 14),
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid.major = element_line(color = "grey30"),
    panel.grid.minor = element_line(color = "grey20"),
    axis.text = element_text(color = "white", size = 12),
    axis.title = element_text(color = "white", size = 14),
    plot.title = element_text(color = "white", size = 18, face = "bold"),
    legend.background = element_rect(fill = "black", color = NA),
    legend.key = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white", size = 12),
    legend.title = element_text(color = "white", size = 14),
    strip.text = element_text(color = "white", size = 14)
  )
```

# Study 1

```{r load data}
# Cleaned data
S1_long_data <- read_csv(here::here("study_1", "data", "clean", "S1_long_data.csv"))
S1_fit_results <-  read_csv(here::here("study_1", "results", "S1_fit_results.csv"))
source(here::here("pre-registered code", "pre-reg_model_implementation.R"))

# Fit Results
S1_fit_results       <- read_csv(here::here("study_1", "results", "S1_fit_results.csv"))
#individual_fits   <- read_csv(here::here("study_1", "results", "S1_individual_fits.csv"))
```

```{r all analyses}
# ----- 2: CREATE DFS FOR WORST PERFOMRERS AND FOR CONDITIONAL PROBABILITIES -----------

# ------ Subset S1_data_long to worst 30% of participants -------
n_participants <- S1_long_data %>% 
  distinct(participant_id) %>% 
  nrow()
n_worst <- floor(n_participants * 0.30)

worst_ids <- S1_long_data %>%
  distinct(participant_id, total_score) %>%
  arrange(total_score) %>%
  slice_head(n = n_worst) %>%
  pull(participant_id)

data_worst_30 <- S1_long_data %>%
  filter(participant_id %in% worst_ids)

#------True and Judged Conditional Probabilities (FULL DATASET)-------

questions <- paste0("q", 1:15, "mark")

# compute true P(q_j=1 | q_i=1) - for success == 1
true_prob_mat_correct <- sapply(questions, function(i) {
  sapply(questions, function(j) {
    mean(S1_long_data[[j]][ S1_long_data[[i]] == 1 ], na.rm=TRUE)
  })
})
diag(true_prob_mat_correct) <- 1
true_df_correct <- as.data.frame(true_prob_mat_correct)
colnames(true_df_correct) <- paste0("given_q", 1:15, "correct")
rownames(true_df_correct) <- paste0("P(q", 1:15, "_correct)")

# compute true P(q_j=1 | q_i=0) - for success == 0
true_prob_mat_incorrect <- sapply(questions, function(i) {
  sapply(questions, function(j) {
    mean(S1_long_data[[j]][ S1_long_data[[i]] == 0 ], na.rm=TRUE)
  })
})
# set diag to NA
diag(true_prob_mat_incorrect) <- NA
true_df_incorrect <- as.data.frame(true_prob_mat_incorrect)
colnames(true_df_incorrect) <- paste0("given_q", 1:15, "incorrect")
rownames(true_df_incorrect) <- paste0("P(q", 1:15, "_correct)")

# create true long by combining correct and incorrect
true_long_correct <- true_df_correct %>%
  as_tibble(rownames="question_evaluated") %>%
  mutate(
    question_evaluated = str_extract(question_evaluated, "\\d+") %>% as.integer(),
    success_observed   = 1
  ) %>%
  pivot_longer(
    cols      = starts_with("given_"),
    names_to  = "given",
    values_to = "true_conditional_prob"
  )

true_long_incorrect <- true_df_incorrect %>%
  as_tibble(rownames="question_evaluated") %>%
  mutate(
    question_evaluated = str_extract(question_evaluated, "\\d+") %>% as.integer(),
    success_observed   = 0
  ) %>%
  pivot_longer(
    cols      = starts_with("given_"),
    names_to  = "given",
    values_to = "true_conditional_prob"
  )

true_long <- bind_rows(true_long_correct, true_long_incorrect)


# create judged long df
judged_long <- S1_long_data %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    average_judged_conditional_prob = mean(human_evaluation == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    given = paste0("given_q", question_observed,
                   if_else(success_observed == 1, "correct", "incorrect"))
  ) %>%
  select(question_evaluated, success_observed, given, average_judged_conditional_prob)

# combine judged and true into a single df
full_prob_df <- inner_join(true_long, judged_long,
                           by = c("question_evaluated","given", "success_observed")) %>%
  filter(question_evaluated != as.integer(str_extract(given, "(?<=given_q)\\d+"))) %>%
  mutate(
    # pull out question_observed
    question_observed = as.integer(str_extract(given, "(?<=given_q)\\d+"))
  ) %>%
  # reorder columns
  select(success_observed, question_observed, everything())


#------True and Judged Conditional Probabilities (WORST 30% DATASET)-------
# repeat process above to create a true and judged conditional probability table for the worst 30% of participants

questions <- paste0("q", 1:15, "mark")

true_mat_worst_correct <- sapply(questions, function(i) {
  sapply(questions, function(j) {
    mean(data_worst_30[[j]][ data_worst_30[[i]] == 1 ], na.rm = TRUE)
  })
})
diag(true_mat_worst_correct) <- 1
true_df_worst_correct <- as.data.frame(true_mat_worst_correct)
colnames(true_df_worst_correct) <- paste0("given_q", 1:15, "correct")
rownames(true_df_worst_correct) <- paste0("P(q", 1:15, "_correct)")

true_mat_worst_incorrect <- sapply(questions, function(i) {
  sapply(questions, function(j) {
    mean(data_worst_30[[j]][ data_worst_30[[i]] == 0 ], na.rm = TRUE)
  })
})
diag(true_mat_worst_incorrect) <- NA
true_df_worst_incorrect <- as.data.frame(true_mat_worst_incorrect)
colnames(true_df_worst_incorrect) <- paste0("given_q", 1:15, "incorrect")
rownames(true_df_worst_incorrect) <- paste0("P(q", 1:15, "_correct)")

true_long_worst <- bind_rows(
  true_df_worst_correct %>%
    as_tibble(rownames = "question_evaluated") %>%
    mutate(
      question_evaluated = str_extract(question_evaluated, "\\d+") %>% as.integer(),
      success_observed   = 1
    ) %>%
    pivot_longer(
      cols      = starts_with("given_q"),
      names_to  = "given",
      values_to = "true_conditional_prob"
    ),
  
  true_df_worst_incorrect %>%
    as_tibble(rownames = "question_evaluated") %>%
    mutate(
      question_evaluated = str_extract(question_evaluated, "\\d+") %>% as.integer(),
      success_observed   = 0
    ) %>%
    pivot_longer(
      cols      = starts_with("given_q"),
      names_to  = "given",
      values_to = "true_conditional_prob"
    )
)

judged_long_worst <- data_worst_30 %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    average_judged_conditional_prob = mean(human_evaluation == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    given = paste0("given_q", question_observed,
                   if_else(success_observed == 1, "correct", "incorrect"))
  ) %>%
  select(question_evaluated, success_observed, given, average_judged_conditional_prob)

worst_30_prob_df <- inner_join(
  true_long_worst,
  judged_long_worst,
  by = c("question_evaluated", "success_observed", "given")
) %>%
  filter(question_evaluated != as.integer(str_extract(given, "(?<=given_q)\\d+"))) %>%
  mutate(
    # pull out question_observed
    question_observed = as.integer(str_extract(given, "(?<=given_q)\\d+"))
  )



rm(judged_long, judged_long_worst, true_df_correct, true_df_incorrect, true_df_worst_correct, true_df_worst_incorrect, true_long, true_long_correct, true_long_incorrect, true_long_worst)
rm(true_mat_worst_correct, true_mat_worst_incorrect, true_prob_mat_correct, true_prob_mat_incorrect)


# ---- CREATE SUMMARIZED DF -----
S1_summarized_data <- S1_long_data %>%
  group_by(success_observed, question_observed, question_evaluated) %>%
  summarise(
    sd_judged_conditional_prob                = sd(human_evaluation, na.rm=TRUE),
    observed_q_perceived_difficulty_unbounded = mean(observed_q_perceived_difficulty_unbounded),
    new_q_perceived_difficulty_unbounded      = mean(new_q_perceived_difficulty_unbounded),
    .groups = "drop"
  ) %>%
  filter(question_observed != question_evaluated)

S1_summarized_data <- S1_summarized_data %>%
  left_join(
    full_prob_df %>%
      select(success_observed, question_observed, question_evaluated,
             average_judged_conditional_prob,
             true_conditional_prob),
    by = c("success_observed","question_observed","question_evaluated")
  )


# ------ EXTRACT RESULTS FROM COMPUTATIONAL MODEL  ----

# ---- Compute BICs ----
S1_MAIN_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$LLMax + 4 * log(nrow(S1_long_data))
S1_NULL_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL", ]$LLMax + 1 * log(nrow(S1_long_data))
S1_NULL_2_BIC <- -2 * S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$LLMax + 2 * log(nrow(S1_long_data))

delta <- 0.005
COMP <- seq(-4, 4, by = delta)

mu <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$mu
sigma <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$sigma
noise <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$noise
beta <- S1_fit_results[S1_fit_results$model == "S1_MAIN", ]$beta

S1_prior_COMP <- dnorm(COMP, mu, sigma)
S1_prior_COMP <- S1_prior_COMP / sum(S1_prior_COMP * delta)


S1_summarized_data$prob_bayes <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  MoreArgs = list(
    prior = S1_prior_COMP, # Pass entire vector as constant
    success = 1,
    noise = noise,
    beta = beta
  )
)

beta <- S1_fit_results[S1_fit_results$model == "S1_NULL", ]$beta
S1_NULL_COMP <- rep(1, length(COMP))
S1_NULL_COMP <- S1_NULL_COMP / sum(S1_NULL_COMP * delta)

S1_summarized_data$prob_null <- mapply(
  FUN = update_and_predict_p_success,
  observed_q_perceived_difficulty_unbounded = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new_q_perceived_difficulty_unbounded = S1_summarized_data$new_q_perceived_difficulty_unbounded,
  MoreArgs = list(
    prior = S1_NULL_COMP,
    success = 1,
    noise = 0,
    beta = beta
  )
)

beta <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$beta
difference <- beta <- S1_fit_results[S1_fit_results$model == "S1_NULL_2", ]$diff

S1_summarized_data$prob_null_2 <- mapply(
  FUN = function(obs, new) {
    soft_max(ifelse(obs + difference >= new, 1, 0), beta = beta)
  },
  obs = S1_summarized_data$observed_q_perceived_difficulty_unbounded,
  new = S1_summarized_data$new_q_perceived_difficulty_unbounded
)

# -------- PRE-REGISTERED ANALYSIS FOR HYPOTHESIS TESTS ----------

model_h1 = lmer(observed_q_objective_difficulty ~ observed_q_perceived_difficulty + (1|participant_id) + (1|question_observed), S1_long_data)
model_h2 <- lm(scale(average_judged_conditional_prob) ~ scale(true_conditional_prob), data = S1_summarized_data)
model_h1m = lmer(observed_q_objective_difficulty ~ observed_q_perceived_difficulty + (1|participant_id) + (1|question_observed), data_worst_30)
model_h2m = lm(average_judged_conditional_prob ~ true_conditional_prob, worst_30_prob_df)
bayes_corr_h3 <- cor.test(S1_summarized_data$average_judged_conditional_prob, S1_summarized_data$prob_bayes)

#for H4: S1_MAIN_BIC < S1_NULL_BIC 
# for H5: S1_MAIN_BIC < S1_NULL_2_BIC

# ----- EXPLORATORY RESEARCH QUESTIONS ---------

model_bayes <- lm(scale(average_judged_conditional_prob) ~ scale(prob_bayes), data = S1_summarized_data)
model_null1 <- lm(scale(average_judged_conditional_prob) ~ scale(prob_null), data = S1_summarized_data)
model_null2 <- lm(scale(average_judged_conditional_prob) ~ scale(prob_null_2), data = S1_summarized_data)


rq1 <- S1_summarized_data %>%
  group_by(success_observed) %>%
  summarise(
    cor = cor(average_judged_conditional_prob, true_conditional_prob, use = "complete.obs"),
    p.value = cor.test(average_judged_conditional_prob, true_conditional_prob)$p.value,
    n.cells = n(),
    .groups = "drop"
  )

# rq2 and rq3 use individual level data (Marius)

# other correlations outside hypotheses
null_cor <- cor.test(S1_summarized_data$average_judged_conditional_prob, S1_summarized_data$prob_null)
null2_cor <- cor.test(S1_summarized_data$average_judged_conditional_prob, S1_summarized_data$prob_null_2)
prob_corr <- cor.test(S1_summarized_data$true_conditional_prob, S1_summarized_data$average_judged_conditional_prob)
diff_corr <- cor.test(S1_long_data$observed_q_perceived_difficulty_unbounded, S1_long_data$observed_q_objective_difficulty_unbounded)

```

## Behavioural Results

**H1: For different questions, participants can accurately assess the rarity of reaching an accurate answer within the time-limit.**

**H3:** The predictions of the Main Model are positively correlated with participants' average predictions. 

**H4:** When fitted to the aggregated dataset, the Main Model better captures participants' behavior than heuristic model 1. 

**H5:** When fitted to the aggregated dataset, the Main Model better captures participants' behavior than heuristic model 2. 

***Exploratory Analysis:***

**RQ1:** If we divide the data by whether the participant observes success or failure of the virtual agent, do participants’ mean predictions of the probabilities of success correlate with the true conditional probabilities? 

2 

**RQ2:** When fitting individual participant-specific models, does the Main Model outperform heuristic model 1? **RQ3:** When fitting individual participant-specific models, does the Main Model outperform heuristic model 2? 

***Manipulation Checks:***

**MC1:** Arithmetic questions are nested. 

**MC1’:** The more difficult the question, the higher the average performance score of the people who got it right (directly derived from nestedness). 
